---
title: "Replication document for Extreme Events Review"
author: "Mathew Stewart"
output: pdf_document
---

## Load Libraries

```{r libraries, message=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
library(tidyverse)
library(vegan)
library(tidytext)
library(gplots)
library(reshape2)
library(here)
library(tm)
```

## Set working directory and load data

```{r setup}
df <- read.csv(here("Appendix_1.csv"))
```

## Examine concepts

The below example examines and plots the most common concepts in the literature sample. To select other variables, use the following in place of "`^4.`":

Variable | Search String |
--- | --- |
Focus | `^2.` |
Definition | `^3.` |
Concept | `^4.` |
Data type | `^5.` |
Analysis type | `^6.` |
Themes | `^7.` |
Geography | `^11.` |
Event type | `^12.` |
Discipline | `^14.` |

```{r concepts}
concept <- df %>%
  separate_rows(Categories, sep = ";") %>%
  mutate(
    Categories = trimws(Categories, which = "left"),
    concept = str_detect(Categories, "^4.")
  ) %>%
  filter(concept == "TRUE") %>%
  count(Categories) %>%
  mutate(percent = n / 200 * 100) %>%
  top_n(10)  %>%
  mutate(Categories = str_replace(Categories, ".*?\\s", ""))

conceptPlot <- ggplot(concept, aes(x = reorder(Categories, +percent), y = percent)) +
  geom_bar(stat = "identity", fill = "#045a8d") +
  theme_classic() +
  xlab("Concept") +
  coord_flip() +
  ylab("%") +
  theme(axis.text = element_text(size = 18)) +
  theme(axis.title = element_text(size = 20, face = "bold")) +
  geom_text(aes(label = n), hjust = 2, colour = "white")

conceptPlot
```

## Examine concepts by discipline

The below example examines and plots the most common concepts in the social sciences. To examine other disciplines change the discipline and denominator to the following:

Discipline | Denominator |
--- | --- |
Earth Science | 38 |
Climate Science | 28 |
Ecology | 23 |
Palaeo Science | 29 |

As above, to examine themes instead of concepts, change the "`^4.`" to a "`^7.`".

```{r discipline}
discipline <- df %>% 
  mutate(discipline = str_detect(Categories, "Social Science")) %>% 
  filter(discipline == "TRUE") %>% 
  separate_rows(Categories, sep = ";") %>% 
  mutate(Categories = trimws(Categories, which = c("left")),
         discipline = str_detect(Categories, "^4.")) %>% 
  filter(discipline == "TRUE") %>% 
  mutate(Categories = str_replace(Categories, "High-magnitude", "Magnitude")) %>% 
  mutate(Categories = str_replace(Categories, ".*?\\s", "")) %>%
  count(Categories) %>%
  mutate(percent = n / 71 * 100) %>%
  slice_max(percent, n = 8, with_ties = TRUE)

socialPlot <- ggplot(discipline, aes(x = reorder(Categories, +percent), y = percent)) +
  geom_bar(stat = "identity", fill = "#045a8d") +
  theme_bw() +
  xlab("") +
  coord_flip() +
  ylab("%") +
  theme(axis.text = element_text(size = 18)) +
  theme(axis.title = element_text(size = 20, face = "bold")) +
  geom_text(aes(label = n), hjust = 2, colour = "white") +
  ggtitle("Social Sciences") +
  theme(plot.title = element_text(face = "bold", size = 22))

socialPlot
```

## Publication dates

Produces two plots. The first shows the number of articles published by year. The second is the same by corrected for through-time changes in publication volume.

```{r publication}
PubDate <- df %>% 
  count(Year.derived) %>%
  mutate(percent = n / 200 * 100) %>% 
  rename(PY = Year.derived) %>% 
  filter(PY > 1995 & PY < 2020)

PubDatePlot <- ggplot(PubDate, aes(x = PY, y = n)) +
  geom_point(size = 5) +
  theme_bw() +
  xlab("Publication year") +
  ylab("Count") +
  theme(axis.text = element_text(size = 18)) +
  theme(axis.title = element_text(size = 20)) +
  xlim(1995, 2020)

PubDatePlot

PubDateWoS <-
  read.csv("Appendix_2.csv") %>% 
  replace(is.na(.), 0) %>%
  select(c(-2,-4)) %>%
  mutate(sum = rowSums(.[2:7])) %>%
  select(c(1,8))

PubDateCor <- merge(PubDateWoS, PubDate, by.x = "PY")

PubDateCorPlot <- PubDateCor %>%
  mutate(percent2 = n / sum * 100) %>%
  ggplot(aes(x = PY, y = percent2)) +
  geom_point(size = 5) +
  theme_bw() +
  xlab("Publication year") +
  ylab("Proportion") +
  theme(axis.text = element_text(size = 18)) +
  theme(axis.title = element_text(size = 20)) +
  ylim(0.008, 0.35) +
  xlim(1997, 2021)

PubDateCorPlot
```

### Collocation (co-occurence) analysis

Analyses the collocation (i.e. co-occurrence) strength between terms. The below example is for the concept "Impact".

To examine the collocation strength of other terms switch out the term "Impact". For example, switch out "Impact" for "Disaster".

As above, to examine themes as opposed to concepts, change the "`^4.`" to a "`^7.`".

```{r collocation}
coll <- df %>% 
  separate_rows(Categories, sep = ";") %>% 
  mutate(Categories = trimws(Categories, which = c("left")),
         coll = str_detect(Categories, "^4.")) %>% 
  filter(coll == "TRUE") %>% 
  group_by(Categories) %>%
  mutate(Categories = str_replace(Categories, ".*?\\s", "")) %>%
  mutate(count = n())

coll_dtm <- coll %>%
  cast_dtm(DOI.name, Categories, count)

w <- dcast(coll, Categories~DOI.name)
x <- as.matrix(w[,-1])
x[is.na(x)] <- 0
x <- apply(x, 2,  function(x) as.numeric(x > 0))
coocCounts <- x %*% t(x) ### create matrix
dimnames(coocCounts) <- list(w[, 1], w[,1])
coocCounts <- as.data.frame(coocCounts)

source("https://slcladal.github.io/rscripts/calculateCoocStatistics.R")

coocTerm <- "Impact"
coocs <- calculateCoocStatistics(coocTerm, coll_dtm, measure = "DICE")

coocdf <- coocs %>%
  as.data.frame() %>%
  dplyr::mutate(CollStrength = coocs,
                Term = names(coocs))
```

## Collectors curves

The below example produces a collectors curve for concepts. To select other variable use the following in place of "^4.":

Variable | Search String |
--- | --- |
Data type | `^5.` |
Theme | `^7.` |
Event type | `^12.` |

```{r collectors}
concept <- df %>% 
  separate_rows(Categories, sep = ";") %>% 
  mutate(Categories = trimws(Categories, which = c("left")),
         concept = str_detect(Categories, "^4.")) %>% 
  filter(concept == "TRUE")

concept <- as.data.frame(with(concept, unclass(table(DOI.name, Categories))))

SAC <- specaccum(concept, method = "random", permutations = 100)
plot(
  SAC,
  xlab = "Articles",
  ylab = "Concepts",
  col = "#045a8d",
  lwd = 1,
  ci.col = "#9ecae1"
)
```

## Cluster analysis

Performs a cluster analysis for the most common themes and produces a heat map.

```{r cluster}
common_themes <- read.csv("Appendix_4.csv") %>%
  separate_rows(Categories, sep = ";") %>%
  mutate(Categories = trimws(Categories, which = c("left")),
         theme = str_detect(Categories, "^7.")) %>%
  filter(theme == "TRUE") %>%
  group_by(Categories) %>%
  mutate(count = n()) %>%
  mutate(Categories = str_replace(Categories, ".*?\\s", "")) %>% 
  filter(count > 18)

as.data.frame(with(common_themes, unclass(table(Author, Categories)))) %>%
  as.matrix() %>%
  heatmap.2(
    key = FALSE,
    trace = "none",
    col = c("white", "black"),
    margins = c(11, 11)
  )
```
